---
title: "Party or Climate? The Polarization of the US Public’s Energy Preferences"
fig-cap-location: top
fontfamily: ebgaramond
link-bibliography: false
format:
  pdf:
    notebook-links: false
    include-in-header:
      text: |
        \usepackage{setspace}
        \usepackage{dcolumn}
        \usepackage{float}
        \usepackage{amsmath}
        \addtokomafont{disposition}{\rmfamily}
  html: 
    include-in-header: 
      text: |
        <style>
          body {
            font-family: 'EB Garamond', serif;
          }
        </style>
  docx: default
---

```{r}
#| include: false
source("scripts/manuscript-setup.R")
```

# Introduction 

Mitigating climate change requires a significant shift away from fossil fuels toward alternative energy sources, such as renewables and nuclear energy. The energy transition requires action from all sectors of society, including governments. Public policy can guide the move away from fossil fuels, yet policy actions require sufficient public support. Public opinion can influence energy policy by incentivizing policymakers through support for clean energy, voting for candidates that advocate for clean energy, and supporting renewable energy projects in their community [@gazmararianPublicOpinionFoundations2026]. Therefore, understanding the mechanisms that shape public opinion is important for the clean energy transition.

It is well-documented that climate change is highly polarized among both political elites and the public, particularly in the United States. Conservatives and Republicans are less likely to think climate change is happening and less supportive of policies to address climate change than Democrats and liberals [@dunlapPoliticalDivideClimate2016; @eganUSPartisanPolarization2024; @hawesClimateSciencePolitics2022]. While views on climate change are extremely polarized among the US public, support for expanding renewable energy sources, such as wind and solar, has been generally strong in the last several years and less polarized than climate change [@croweEffectPartisanCues2020; @ter-mkrtchyanPortionsPortfoliosUnderstanding2022].

Yet, this may be changing, with support for renewable energy becoming more polarized. Recent survey data from the Pew Research Center show that support for expanding renewable energy among Republicans has declined, from 65 percent in 2020 to 28 percent in 2026, while at the same time, support for expanding fossil fuels increased from 35 percent in 2020 to 71 percent in 2026 among Republicans [@kennedyAmericansShiftingViews2026]. In this paper, I examine two possible mechanisms that could account for the shift among Republicans, including framing that connects renewable energy to climate change or cues from prominent Republicans, most notably Donald Trump.

# Cues and Polarization on Energy

There is clear evidence of polarization in views on climate change among the US public, with multiple studies finding stark differences in climate views across political beliefs such as ideology and partisanship [e.g., @druckmanEvidenceMotivatedReasoning2019; @dunlapPoliticalDivideClimate2016; @dunlapWideningGapRepublican2008; @mccrightPoliticizationClimateChange2011; @smithPolarisationClimateEnvironmental2024]. However, political polarization regarding energy, including the favorability of specific energy sources and support for energy policies, is more nuanced [@ansolabehereCheapCleanHow2016]. In general, polarization varies by energy source, with views on fossil fuels being extremely polarized, whereas views on other energy sources are less polarized [@bergquistEnergyPolicyPublic2020].

Public opinion regarding fossil fuels, such as coal, oil, and gas, is shaped, in large part, by political beliefs. Compared to conservatives and Republicans, liberals and Democrats are less supportive of coal [@hartPoliticalPolarizationSupport2021] and more concerned about extraction processes associated with oil and gas, such as hydraulic fracturing, or "fracking" [@davisEnergyAbundanceEnvironmental2014; @raimiPartisanshipProximityPredict2020]. For liberals and Democrats, the risks of fossil fuels, including the health, environmental, and climate impacts, influence opposition to fossil fuels, whereas conservatives and Republicans tend to emphasize the economic and energy security benefits [@clarkePoliticsScientificConsensus2019; @howellSeeingRiskColoredGlasses2019].

Renewable energy sources are generally less polarized than fossil fuels, with liberals/Democrats and conservatives/Republicans supporting increases in renewables relative to the current energy mix [@hawesClimateSciencePolitics2022; @ter-mkrtchyanPortionsPortfoliosUnderstanding2022]. However, there are partisan differences, but they are differences in *degree* rather than *direction*, as liberals/Democrats tend to be more supportive of renewables than conservatives/Republicans [@bergquistEnergyPolicyPublic2020; @hamiltonChangeWindUS2018]. Additionally, the reasons for support vary by political beliefs. Conservatives/Republicans support renewable sources because of the economic benefits and costs, whereas liberals/Democrats support them because of climate concerns [@gustafsonRepublicansDemocratsDiffer2020]. In general, previous research suggests that the way that political beliefs shape the public's views on fossil fuels and renewables is through the emphasis that conservatives/Republicans vs liberals/Democrats place on various aspect of energy. For conservatives/Republicans, the emphasis is on costs and benefits, but for liberals/Democrats the emphasis is on health, environment, and climate change.

Views on nuclear energy are less polarized across political beliefs than renewables sources, but less polarized than fossil fuels. To the extent nuclear energy is polarized, it tends to find more support among conservatives/Republicans than liberals/Democrats [@yeoPartisanAmplificationRisk2014; @mcbethPoliticalIdeologyNuclear2023a]. However, support for nuclear energy is also driven by events, like the three major nuclear accidents, and shifts in the supply of other energy sources [@guptaTrackingNuclearMood2019].

As noted, recent public surveys indicate that support among Republicans for renewables is falling while support for fossil fuels is increasing. One reason for this trend may be that renewable energy has become entangled with climate change, leading Republicans to be less supportive. Therefore, if renewables are framed as an approach to address climate change, they are likely to become more polarized. Framing entails focusing on one or more aspects of an issue, which can shape how that issue is viewed and understood. How climate policies are framed, such as science, moral, and economic frames can lead to more (less) support for those policies[@seversonMoralFramesClimate2015], and frames regarding reducing energy waste and producing low-carbon domestic energy can reduce climate change polarization [@whitmarshToolsNewClimate2017]. Similarly, frames can shape views on energy sources, as a frame that emphasizes how renewables can save costs increases support for renewables across partisan groups [@gustafsonDurableBipartisanEffects2022].

The importance of framing in shaping public opinion in studies that have found that a climate change frame can influence views on energy. For example, one study found that Republicans were less supportive of a carbon tax, fuel efficiency, and renewables when they were framed as a way to address climate change compared to being framed as a way to address air pollution or energy security [@feldmanClimateChangePolarizing2018]. Additionally, Republicans were more supportive of renewable energy portfolios in their state when they were framed as helping with air pollution, but no difference in support when those policies were associated with climate change [@stokesRenewableEnergyPolicy2017]. Along similar lines, views on climate change mediated views about the preferred US energy mix, but only for conservative Republicans and liberal Democrats, not for more moderate political beliefs [@hawesClimateSciencePolitics2022]. However, it should be noted that framing may have limited impacts [@hartPoliticalPolarizationSupport2021; @zhouBoomerangsJavelinsHow2016], particularly in the presence of counter-frames [@aklinDebatingCleanEnergy2013].

Apart from framing, elite cues can drive polarization among the public. Elites such as prominent elected officials, political parties, and media figures can influence public opinion and drive polarization. Indeed, elite cues may be one of the most important factors for public opinion about climate change [@brulleShiftingPublicOpinion2012], and it is likely that the elite debate over the Kyoto Protocol in 1998-99 was a major factor in the development of polarization over climate change [@krosnickImpactFall19972000]. Additionally, cues work to both strengthen the views of co-partisans and produce a backlash effect for out-partisans, which can lead to polarization [@merkleyPartyCuesNews2021]. By taking a position on an issue, elites can bring attention to it and make it more salient, which creates a signal for the public. This signal can lead to polarization, such as with the Green New Deal that was widely popular, but became polarized as a result of elite discourse [@gustafsonDevelopmentPartisanPolarization2019].

Polarization on energy among the public varies by source, yet polarization among elites, specifically elected officials, grew as energy policy became more regulatory and less distributive [@lowryDisentanglingEnergyPolicy2008]. But, it was the 2010s when differences on energy issues between the political parties in the US Congress became "pervasive and intense" [@jeongPolarisationEnergyPolicy2021, 4]. The growing polarization on energy issues among elites sets the stage for more uniform polarization among the public. This becomes more likely as prominent political figures express views on energy issues, as cues that came from individual elites seem to be more impactful on public opinion than cues that are simply based on political party [@nicholsonPolarizingCues2012].

No political figure in the last decade, or more, has done more to polarization the public than President Trump. The "Trump effect" has been show to polarize views on immigration [@taylorTrumpEffectImmigration2026] and free trade [@essigTrumpEffectPolitical2021]. President Trump expressed skepticism about climate change even before running for the presidency. As a political candidate in 2016, he expressed support for fossil fuels, and during his first term, he withdrew the US from the Paris Agreement on climate change. However, since the start of his second administration, President Trump has accelerated efforts to bolster the production and use of fossil fuels while also hindering the development of renewable energy sources [@gerrardTrumpsMultiProngedAttack2025; @joselowHowTrumpsFirst2025]. Such actions are likely to polarize the public further on energy issues.

The observed growing polarization on energy issues, most notably fossil fuels and renewables, could be a result of a climate change frame associated with those energy sources and/or a result of elite cues, particularly from President Trump, regarding those energy sources. In the next section, I describe the research design I used to examine the impacts of a climate change cue or a Trump cue on the US public's preferred energy mix.

# Research Design

The purpose of the study is to examine how partisan or climate change cues could shift the general public's preferred energy mix for electricity production. I use a 3 X 2 experimental design to test several hypotheses regarding the impact of cues on the shifts in the preferred energy mix for producing electricity across political beliefs. Respondents were randomly assigned to one of three experimental conditions, including a Trump cue, a climate cue, and a control condition with no cues.

The preferred energy mix is main the dependent variable, and is measured as the percentage each respondent allocated across the five energy sources, including fossil fuels, wind, solar, hydro, and nuclear. As part of the research design, respondents were first given the current energy mix of the US related to electricity, which is that the US obtains 58 percent of our electricity from fossil fuels (coal, oil, and natural gas), 22 percent from renewable sources (wind, solar, and hydropower), 18 percent from nuclear energy, and 2 percent from other sources (wood, biofuels, waste products, and geothermal energy). Therefore, respondents were anchored to the actual energy mix use to produce electricity in the US, as of 2024 the most recent year where data was available. Following the information, respondents were place in one of three conditions, and then asked for their preferred energy mix.

## Hypotheses

The research design is intended to test the following hypotheses, which were preregistered using the Open Science Framework (OSF)[^1]

[^1]: See here: https://osf.io/s2ku8/overview?view_only=8d073e8b77f44768a27184b9b0aff785

> Conservative Republicans who receive the Trump cue will support a *higher* percentage of fossil fuels than those in the climate change or control group, and those with other political beliefs

> Conservative Republicans who receive the climate cue will support a *higher* percentage of fossil fuels than those in the Trump or control group, and those with other political beliefs

> Conservative Republicans who receive the Trump cue will support a *lower* percentage of renewable energy sources than those in the climate change or control group, and those with other political beliefs

> Conservative Republicans who receive the climate cue will support a *lower* percentage of renewable energy sources than those in the climate change or control group, and those with other political beliefs

> Liberal Democrats who receive the Trump cue will support a *lower* percentage of fossil fuels than those in the climate change or control group, and those with other political beliefs

> Liberal Democrats who receive the Trump cue will support a *higher* percentage of renewable energy sources than those in the climate change or control group, and those with other political beliefs

> Liberal Democrats who receive the climate cue will support a *lower* percentage of fossil fuels than those in the climate change or control group, and those with other political beliefs

> Liberal Democrats who receive the climate cue will support a *higher* percentage of renewable energy sources than those in the climate change or control group, and those with other political beliefs

## Data

The data for this study comes from a survey of `r n_total` US residents recruited through Cloud Research, a survey panel provider that uses a variety of online recruitment methods to build a large and diverse panel of respondents. Cloud Research provides a quota-based sampling approach, where US Census data is used to ensure representation across key demographic groups. The survey was internet based, and administered in QuestionPro. It was fielded from April 10 to April 15, 2026. Survey weights were applied to ensure that the sample was representative.[^3] 

As noted, respondents were randomly assigned to one of three conditions, including a Trump cue, a climate change cue, or a control condition with no cues. The Trump cue was in the form of a question regarding agreement, on a 1 strongly disagree to 5 strongly agree scale, with Trump's actions on encouraging the development of fossil fuels over renewable energy. The climate cue was a question regarding agreement with the need to increase the use of renewable sources to address climate change. Random assignment produced approximately equal groups, with the Trump cue, *n* = `r n_trump`, the climate cue, *n* = `r n_climate`, and control, *n* = `r n_control`, from the sample of *N* = `r n_total`).[^4]



[^3]: For more information on the sample and the weighting procedures see the online methods section.

[^4]: A comparison of covariates across the three conditions is presented in the supplemental methods appendix.

## Measures

As noted, the main dependent variable is the preferred energy mix, where respondents indicated their preferred percentages for each source. It was worded as follows:

> Thinking about the overall supply of electricity. What percentage of the total U.S. electricity supply over the next 20 years would you like to see come from each of the 5 primary sources?

> Fossil fuels (coal, oil, natural gas), which produced 58 percent in 2024 Wind energy, which produced 10 percent in 2024 Solar energy, which produced 7 percent in 2024 Hydropower, which produced 5 percent in 2024 Nuclear energy, which produced 18 percent in 2024

The Trump and climate cues were questions that asked respondents to indicate their agreement with the statements on a 1 (strongly oppose) to 5 (strongly support) scale. Question wording included:

> Trump cue: *As you may have heard, the Trump administration has taken steps to increase the production of fossil fuels, particularly coal, and reduce the use of renewable energy like solar and wind for electricity generation*

> Climate cue: *Scientists agree that the use of fossil fuels has contributed to climate change, and to address climate change, we need to increase the use of renewable sources such as wind, solar, and hydropower*

To measure political beliefs, respondents were asked with which political party they most identify and to place themselves on a 1 to 7 scale of political ideology, where 1 is strongly liberal, 4 is middle of the road, and 7 is strongly conservative. Response were combined so that liberal Democrats are those that identify with the Democratic party and are strongly liberal or liberal, and conservative Republicans are those that are Republicans and strongly conservative or conservative. Other political beliefs, including conservative Democrats, liberal Republicans, and moderate independents are the excluded referents.

In the next section, I first examine the cue questions to confirm that respondents were responsive to the cues. Then, I examine the preferred energy mix by political beliefs and treatment condition. Next, I test whether the partisan gaps in preferred energy mix differ significantly across treatment conditions using a difference-in-differences test. Finally, I explore the differences in preferred energy sources from the current energy mix across treatment conditions and political beliefs.

# Results

I first use OLS models to estimate agreement with the Trump and climate cue questions for liberal Democrats and conservative Republicans. @fig-cueModels shows the coefficient plots.

```{r}
#| label: fig-cueModels
#| fig-cap: Coefficient Plots of Trump Cue and Climate Cue
#| echo: false
#| warning: false
#| message: false

# coefs comes from scripts/manuscript-setup.R (sourced in cue-energy.qmd)
p1 <- ggplot(filter(coefs, model == "Trump Cue"), aes(x = estimate, y = term, color = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(size = 4) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_manual(values = c("Conservative\nRepublicans" = "red",
                                "Liberal\nDemocrats" = "blue")) +
  labs(x = "", y = NULL,
       title = "How do you feel about the Trump administration’s goal of increasing the use of\nfossil fuels and decreasing the use of renewable sources in generating electricity?") +
  theme_minimal() + theme(legend.position = "none", plot.title = element_text(face = "bold", size = 8))

p2 <- ggplot(filter(coefs, model == "Climate Cue"), aes(x = estimate, y = term, color = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(size = 4) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_manual(values = c("Conservative\nRepublicans" = "red",
                                "Liberal\nDemocrats" = "blue")) +
  labs(x = "Coefficient Estimates", y = NULL,
       title = "How do you feel about increasing the use of renewable sources\nin generating electricity as a way to address climate change?",
       caption = "A 1 to 5 scale was used on both questions") +
  theme_minimal() + theme(legend.position = "none", plot.title = element_text(face = "bold", size = 8))

p1 / p2
```

As shown, for the Trump cue liberal democrats were significantly less likely, at `r round(cuemodel1$coefficients[3], 3)`, to support Trump's actions on energy, whereas conservative Republicans had an estimated level of support `r round(cuemodel1$coefficients[2], 3)` points higher on the 1 to 5 scale than other political beliefs. For the climate change cue, liberal Democrats were significantly more likely than others to support renewable energy to address climate change, with an estimated coefficient of `r round(cuemodel2$coefficients[3], 3)`, and conservative Republicans were significantly less likely with a coefficient of `r round(cuemodel2$coefficients[2], 3)`. These results suggest that respondents were responsive to the cues in ways that were expected.

Next, respondents were asked what percentage of the total US electricity supply should come from various sources. Respondents were again reminded of the current energy mix. Additionally, respondents were instructed that their percentages for each source must add to one hundred. The results by beliefs and cue are shown in @fig-pred.

```{r}
#| label: fig-pred
#| fig-cap: Predicted Energy Mix by Political Beliefs and Cues
#| echo: false
#| warning: false
#| message: false

# pred_df comes from scripts/manuscript-setup.R (sourced in cue-energy.qmd)
ggplot(pred_df, aes(x = treatment, y = fit, fill = party)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(
    aes(ymin = lwr, ymax = upr),
    position = position_dodge(width = 0.8),
    width = 0.2
  ) +
  facet_wrap(~ source, nrow = 1) +
  scale_fill_manual(values = c("Liberal Democrats" = "#2166AC",
                               "Conservative Republicans" = "#B2182B")) +
  labs(
    x = NULL,
    y = "Percent (%)",
    fill = NULL,
    title = "",
    caption = "Bars show predicted values from OLS; whiskers are 95% confidence intervals."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(size = 5, hjust = 0.5),
    axis.title = element_text(size = 10),
    panel.grid.major.x = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

As can be seen, conservative Republicans preferred an energy mix with more fossil fuels than did liberal Democrats, particularly Republicans that received the Trump cue. Conservative Republicans in the Trump condition preferred an energy mix that consists of about `r round(pred_df$fit[2],0)`% fossil fuels compared to `r round(pred_df$fit[12],0)`% in the climate condition and `r round(pred_df$fit[22],0)`% in the control. Liberal Democrats preferred an energy mix with more renewables than Republicans, particularly wind and solar, and the percentages of renewable sources seemed consistent for Democrats across the three conditions. Finally, Republicans preferred more nuclear energy than Democrats across all conditions, yet the differences were smallest in the Trump condition. 

To assess which of the differences in @fig-pred are statistically significant, I estimated design-based contrasts both for the partisan gap (liberal Democrats − conservative Republicans) within each condition and for shifts across conditions (e.g., Trump − control). The results are shown in @tbl-pred-contrasts. Additionally, I illustrate the polarization of energy preferences across beliefs and cues by sorting the gaps by size and direction, and they are shown in @fig-gap.  

```{r}
#| label: tbl-pred-contrasts
#| tbl-cap: "Pairwise Differences in the Predicted Preferred Energy Mix"
#| echo: false
#| warning: false
#| message: false
#| tbl-pos: H

make_pred_contrasts_flextable()
```


```{r}
#| label: fig-gap
#| fig-cap: Largest Gaps by Political Belief in Preferred Energy Mix
#| echo: false
#| warning: false
#| message: false

# gap_sorted / gap_ranked come from scripts/manuscript-setup.R (sourced in cue-energy.qmd)
ggplot(gap_sorted, aes(x = gap, y = label)) +
  geom_vline(xintercept = 0, color = "grey40", linewidth = 0.4) +
  geom_segment(aes(x = 0, xend = gap, yend = label),
               color = "grey60", linewidth = 0.6) +
  geom_pointrange(aes(xmin = lwr, xmax = upr, color = direction), size = 0.5) +
  scale_color_manual(
    values = c("Dem" = "#2166AC", "Rep" = "#B2182B"),
    labels = c("Dem" = "Democrats prefer more",
               "Rep" = "Republicans prefer more"),
    name = NULL
  ) +
  labs(
    x = "Predicted partisan gap (percentage points)\nLiberal Democrats − Conservative Republicans",
    y = NULL,
    title = "",
    caption = "Sorted by absolute size of gap. Bars are 95% confidence intervals from svyglm models."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```

The partisan gap is statistically significant for every energy source in every condition, which means that liberal Democrats and conservative Republicans differ in their preferred mix regardless of the cue. As shown in @fig-gap, fossil fuels are the most polarized, particularly in the Trump condition with an average difference of `r round(abs(gap_ranked$gap[1]),0)` percentage points between liberal Democrats and conservative Republicans. In general, the within-party cue effects are concentrated almost entirely among conservative Republicans exposed to the Trump cue. 

Relative to the control condition, conservative Republicans in the Trump condition preferred `r pred_diff("Fossil Fuels", "Republicans: Trump - Control", abs = TRUE)` more fossil fuels, and less renewables with `r pred_diff("Wind", "Republicans: Trump - Control", abs = TRUE)` less wind, `r pred_diff("Solar", "Republicans: Trump - Control", abs = TRUE)` less solar, and `r pred_diff("Hydro", "Republicans: Trump - Control", abs = TRUE)` less hydro power. By contrast, the climate cue moved Republicans negligibly on fossil fuels, `r pred_diff("Fossil Fuels", "Republicans: Climate - Control")`. 

Also as shown, Democrats preferred more wind and solar energy overall, with the largest partisan gaps in the Trump condition. Liberal Democrats did not shift as much as conservative Republicans across conditions on any source; however, they did, unexpectedly, show some responsiveness to the Trump cues. Compared to the climate condition, liberal Democrats in the Trump condition preferred `r pred_diff("Fossil Fuels", "Democrats: Trump - Climate", abs = TRUE)`more fossil fuels, `r pred_diff("Wind", "Democrats: Trump - Climate", abs = TRUE)`  less wind, and `r pred_diff("Hydro", "Democrats: Trump - Climate", abs = TRUE)` less hydro. Similarly, compared to the control condition, liberal Democrats in the Trump condition preferred more fossil fuels, `r pred_diff("Fossil Fuels", "Democrats: Trump - Control", abs = TRUE)` and `r pred_diff("Solar", "Democrats: Trump - Control", abs = TRUE)` less hydro. However, the shifts for Democrats were much smaller than those for Republicans, and the shifts for Democrats in the Trump condition compared tp the control were significant at p < 0.10.

Nuclear was supported more by conservative Republicans, and was most polarized in the climate and control conditions. No within-party difference across the conditions were significant for nuclear energy. Overall, hydro was preferred by liberal Democrats, yet it was the least polarized energy source. The hydro power control condition had the smallest gap, with a difference of only about `r round(abs(gap_ranked$gap[nrow(gap_ranked)]),0)` percentage points. Liberal Democrats in the climate cue preferred `r pred_diff("Hydro", "Democrats: Climate - Control", abs = TRUE)` more hydro than liberal Democrats in the control condition. Together, these results show that the cue-driven movement in @fig-pred is driven largely by conservative Republicans responding to the Trump cue; however, liberal Democrats also responded to the Trump cue.

Next, I compare the gaps across treatment conditions to examine whether the cues shifted the partisan differences in preferred energy mix. Specifically, I used a difference-in-differences (DiD) approach, where I examine the difference between political beliefs (i.e., liberal Democrats - conservative Republicans), which are shown in @fig-gap, and differences between treatment conditions (e.g., Trump cue - control) using interaction terms between political beliefs and treatment conditions.[^5] This approach allows me to test whether the cues shifted the partisan gaps in preferred energy mix. The results are shown in @tbl-did.

[^5]: Note that I am not using a traditional DiD design, which estimates differences between groups across two time periods based on an assumption of parallel trends [@bakerDifferenceinDifferencesDesignsPractitioners2026].

```{r}
#| label: tbl-did
#| tbl-cap: "Treatment Effects on the Preferred Energy Mix Across Political Beliefs"
#| echo: false
#| warning: false
#| message: false

# did_table and make_did_flextable() come from scripts/manuscript-setup.R
# (sourced in cue-energy.qmd). Rendered in-document so it appears in html, pdf, docx.
make_did_flextable()
```

@tbl-did illustrates the change in the partisan gap for each energy source across each pairwise treatment comparison. Additionally, @tbl-did includes an omnibus F-test which examines whether the cues had an overall effect on the partisan gaps across each energy source. As can be seen, for fossil fuels the F-test was significant, indicating that the partisan gaps in preferred energy mix shift significantly for fossil fuels. These impacts were a result of the Trump condition, as a shift of `r did_shift("Fossil Fuels", "Trump Cue vs Control")` occurred between the Trump and control condition, and a shift of `r did_shift("Fossil Fuels", "Trump Cue vs Climate")` happened between the Trump and climate change condition. Additionally, the gap for hydro power was significantly higher (at *p*\<0.10) in the climate condition compared to the control condition.

Finally, I compared the preferred energy mix to the current baseline. Overall, the evidence suggests that Trump may polarize energy preferences by shifting conservative Republicans to be more in favor of fossil fuels than they might be otherwise. However, the shift in the energy mix from the current baseline is important to address climate change. @fig-mixchange shows the differences in the current vs preferred energy mixes across political beliefs and treatment conditions.

```{r}
#| label: fig-mixchange
#| fig-cap: Preferred Change from Existing Energy Mix (percentage points)
#| echo: false
#| warning: false
#| message: false

# change_df comes from scripts/manuscript-setup.R (sourced in cue-energy.qmd)
ggplot(change_df, aes(x = treatment, y = change_fit, fill = party)) +
  geom_hline(yintercept = 0, color = "grey30", linewidth = 0.5) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(
    aes(ymin = change_lwr, ymax = change_upr),
    position = position_dodge(width = 0.8),
    width = 0.2
  ) +
  facet_wrap(~ source, nrow = 1) +
  scale_fill_manual(values = c("Liberal Democrats"        = "#2166AC",
                               "Conservative Republicans" = "#B2182B")) +
  labs(
    x = NULL,
    y = "Percentage Points",
    fill = NULL,
    title = "",
    caption = paste(
      "Bars show predicted preferred share minus the current share shown to respondents.",
      "Whiskers are 95% confidence intervals from svyglm models with survey weights.",
      "Positive values = expand; negative = contract.",
      sep = "\n"
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(size = 5, hjust = 0.5),
    axis.title = element_text(size = 10),
    panel.grid.major.x = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

As shown, all respondents want an energy mix with fewer fossil fuels and more renewables. When comparing each group's predicted preferred share to the status quo shown to respondents, the preferred changes are statistically significant for 29 of the 30 group-by-source-by-condition comparisons
Both liberal Democrats and conservative Republicans preferred significantly less fossil fuel and significantly more wind, solar, and hydropower than the current mix across all three conditions, with Democrats favoring larger shifts away from fossil fuels and toward renewables than Republicans. These findings support previous research that suggests that the public is generally supportive of a transition to renewable energy sources, even among conservative Republicans. Additionally, the cues had little effect on the preferred change from the current mix for most sources, except for fossil fuels, where the Trump cue led conservative Republicans to prefer a smaller reduction in fossil fuels, `r round(change_df$change_fit[change_df$source == "Fossil Fuels" & change_df$party == "Conservative Republicans" & change_df$treatment == "Trump"], 1)`,  than they did in the control with `r round(change_df$change_fit[change_df$source == "Fossil Fuels" & change_df$party == "Conservative Republicans" & change_df$treatment == "Control"], 1)`, and climate conditions at `r round(change_df$change_fit[change_df$source == "Fossil Fuels" & change_df$party == "Conservative Republicans" & change_df$treatment == "Climate"], 1)`.

The two parties diverge only on nuclear energy, with liberal Democrats preferring a small but significant reduction (about three percentage points), whereas conservative Republicans preferred a significant expansion in the control and climate conditions. The only non-significant difference is conservative Republicans' preferred nuclear share in the Trump condition, which is statistically indistinguishable from the current level (a change of `r round(change_df$change_fit[change_df$source == "Nuclear" & change_df$party == "Conservative Republicans" & change_df$treatment == "Trump"], 1)` percentage points, with a 95% confidence interval that includes zero). 

# Discussion 

It was expected that the cue would influence political beliefs

 - First difference = the partisan gap (Liberal Democrats − Conservative Republicans) within a condition. That's fig-gap.
  - Second difference = how that gap changes when you move from one condition to another. That's the party:treatment interaction coefficient, which is exactly
  what tbl-did reports.


hypotheses only confirmed for fossil fuels; contrary to expectations, the cues had little effect for other energy sources

| Source       | Significance        | Notes                                          |
|--------------|---------------------|------------------------------------------------|
| Fossil Fuels | p ≈ 1e-23 to 1e-27  | Strongest gaps (16–24 pts), Republicans prefer more |
| Wind         | p ≈ 1e-20 to 1e-35  | Democrats prefer more (~11–12 pts)             |
| Solar        | p ≈ 1e-15 to 1e-22  | Democrats prefer more (~10–12 pts)             |
| Nuclear      | p ≈ 1e-5 to 0.008   | Republicans prefer more (4–8 pts)              |
| Hydro        | p ≈ 0.02 to 2e-6    | Weakest gaps (2–4 pts), but still significant  |


\newpage


# References