---
title: "Supplemental Materials: Party or Climate? The Polarization of the US Public’s Energy Preferences"
fig-cap-location: top
fontfamily: ebgaramond
link-bibliography: false
---

```{r}
#| include: false
source("scripts/manuscript-setup.R")
```

# Methods

## Survey Sample

Despite the use of quotas, the sample from Cloud Research was not perfectly representative of the US population. Specifically, the sample was younger, whiter, more educated, and more aligned with the Democratic party than the US population. To address this, I used the `survey` and `anesrake` packages in `R` to perform iterative raking to compute survey weights that align the sample with the US Census's 2023 American Community Survey (ACS) on key demographic variables, including age, race, gender, and education. Additionally, I used the Pew Research Center's National Public Opinion Reference Survey (NPORS) from 2025 to align the sample on political party affiliation.[^1] The differences between the unweighted and weighted sample are shown in @tbl-descript. The weights were then applied to all analyses in the paper.

[^1]: See here:https://www.pewresearch.org/politics/fact-sheet/party-affiliation-fact-sheet-npors/.


```{r create dataset for descriptive statistics}
#| echo: false
#| include: false
#| warning: false

vars_comp <- c("age", "male", "white", "college", "democrat", "republican")

# Friendly labels for the table — same length and order as vars_comp
var_labels <- c(
  age        = "Age (years)",
  male       = "Male",
  white      = "White (non-Hispanic)",
  college    = "College graduate",
  democrat   = "Democrat",
  republican = "Republican"
)

# Build the comparison table
comparison <- data.frame(
  Variable   = var_labels[vars_comp],
  Unweighted = sapply(vars_comp, function(v) mean(cueEnergyData[[v]], na.rm = TRUE)),
  Weighted   = sapply(vars_comp, function(v) {
                 as.numeric(svymean(as.formula(paste0("~", v)),
                                    design, na.rm = TRUE))
               }),
  row.names = NULL
) |>
  mutate(Difference = Weighted - Unweighted)

```

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: tbl-descript
#| tbl-cap: "Unweighted vs. Weighted Means on Key Variables"
#| tbl-pos: H

comparison |>
  mutate(across(c(Weighted, Unweighted, Difference), \(x) round(x, 3))) |>
  flextable() |>
  set_header_labels(Variable = "") |>
  align(j = 1,   align = "left",  part = "all") |>
  align(j = 2:4, align = "right", part = "all") |>
  bold(part = "header") |>
  fontsize(size = 10, part = "all") |>
  padding(padding = 4, part = "all") |>
  autofit() |>
  add_footer_lines(values = "Note: Weights computed via iterative raking to ACS 2023 and Pew NPORS targets.") |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")
```

## Covariate Balance Across Conditions

Respondents were randomly assigned to the Trump cue, climate cue, or control condition. @tbl-balance compares demographic covariates across the three conditions to confirm that random assignment produced comparable groups. Of the covariates examined, the conditions differ significantly only on age and the share of white respondents; however, political beliefs (liberal Democrats and conservative Republicans) are balanced across conditions.

```{r}
#| label: tbl-balance
#| tbl-cap: "Covariate Balance Across Experimental Conditions"
#| echo: false
#| warning: false
#| message: false
#| tbl-pos: H

make_balance_flextable()
```

## Robustness: Controlling for Demographic Covariates

Because the conditions differ by chance on age and race (see @tbl-balance), I re-estimate the main difference-in-differences models adding age, sex, race, education, and income as additive controls. @tbl-did-controls reports the same contrasts as the main-text difference-in-differences table. The estimated treatment effects on the partisan gap are substantively unchanged from the main specification, indicating that the cue effects are not driven by the chance covariate imbalances across conditions.

```{r}
#| label: tbl-did-controls
#| tbl-cap: "Difference-in-Differences Estimates with Demographic Controls"
#| echo: false
#| warning: false
#| message: false
#| tbl-pos: H

make_did_controls_flextable()
```

## Survey Question Wording

Trump cue: As you may have heard, the Trump administration has taken steps to increase the production of fossil fuels, particularly coal, and reduce the use of renewable energy like solar and wind for electricity generation.

Using a scale of 1 to 5, where 1 means strongly oppose and 5 means strongly support, how do you feel about the Trump administration’s goal of increasing the use of fossil fuels and decreasing the use of renewable sources in generating electricity?

Climate cue: Scientists agree that the use of fossil fuels has contributed to climate change, and to address climate change, we need to increase the use of renewable sources such as wind, solar, and hydropower.

Using a scale of 1 to 5, where 1 means strongly disagree and 5 means strongly agree, how do you feel about increasing the use of renewable sources in generating electricity as a way to address climate change?

Thinking about the overall supply of electricity. What percentage of the total U.S. electricity supply over the next 20 years would you like to see come from each of the 5 primary sources?

NOTE: The sum of the five boxes below must equal 100

Fossil fuels (coal, oil, natural gas), which produced 58 percent in 2024 Wind energy, which produced 10 percent in 2024 Solar energy, which produced 7 percent in 2024 Hydropower, which produced 5 percent in 2024 Nuclear energy, which produced 18 percent in 2024

## Preregistered Hypotheses

In addition to the hypotheses present in the paper, I also preregistered other hypotheses that were not included in the paper. These include hypotheses about the effects of political beliefs on support/opposition to the climate change and Trump cues, as well as hypotheses about how the general impacts of the climate change and Trump cues, regardless of political ideology.

CUES Conservative Republicans will be more supportive of Trump's actions to increase the use of fossil fuels and decrease the use of renewable sources than those with other political beliefs

Liberal Democrats will be more supportive of increasing the use of renewable sources to address climate change than those with other political beliefs

ENERGY MIX Those who receive the Trump cue will support a higher percentage of fossil fuels than those in the climate change or control group

Those who receive the climate cue will support a higher percentage of renewable energy sources (wind, solar, hydropower) than those in the Trump or control group